{\color{red}\textbf{TODO}}

### 2024-12-15 Paper draft/draft.tex
Updated schemes

\title{Draft of paper}

@@

\address[First]{University di Trento, Trento, Italy (e-mail: marco.sterlini@studenti.unitn.it) }
\address[Second]{LAAS-CNRS, Université de Toulouse, CNRS, Toulouse, France (e-mail:bu) }

\begin{abstract}
\textbf{INSERT ABSTRACT}
\end{abstract}

\begin{keyword}
\textbf{INSERT KEYWORDS}
\end{keyword}

@@

\section{Introduction}
\textbf{INSERT INTRODUCTION}\\
\emph{Notation:} $\NN, \RR^{n}, \RR^{n \times m}$ denote the set of natural non-negative integers, the set of real vectors of dimension $n$ and the set of real matrices of dimension $n \times m$, respectively. For any matrix $A$, $A\tr$ is its transpose. For any square matrix $A$, the operator $\He{ A } = A + A\tr$ is defined. $\texttt{diag}(A_1, A_2)$ is a block-diagonal matrix with block-diagonal matrices $A_1$ and $A_2$. For a partitioned matrix, the symbol $\star$ stands for symmetric blocks. $\1_n = \bmx{1 & \dots & 1}\tr \in \NN^{n \times 1}$. We identify with subscript $i$ the $i$-th element of a vector or the $i$-th row of a matrix.

\section{Problem Formulation}
We consider the non-linear discrete-time system stabilized by a Neural Network (NN) controller represented in figure \ref{fig:first_scheme} described by:
\begin{equation}
  x^{+} = A x + B \texttt{sat}(\bar{u}) + C \Phi(E x) + D d 
\end{equation}
where the saturation function \texttt{sat} is defined as:
$$
    \texttt{sat}(\bar{u}) = \texttt{sign}(\bar{u})\text{min}(|\bar{u}|, \hat{u})
$$
with $\hat{u}$ the limit of the input signal, $x \in \RR^{n_x}$ the state vector, $\bar{u} \in \RR^{n_u}$ the raw output of the controller, $d \in \RR^{n_d}$ a constant disturbance input  and $\Phi \in \RR^{n_q}$ the representation of the non-linearity of the system. 
\begin{assumption}\label{ass:nonlin}
We assume that $\Phi$ satisfies some local quadratic abstraction for some $\underline{\mu}, \bar{\mu} > 0 \in \RR^{n_q}, \bar{S}, \bar{Q}, \bar{W} \in \RR^{n_q \times n_q}$
\begin{equation}\label{eqn:nonlin-set}
\forall v \in \bar{S} = \left\{ v \in  \RR^{n_q} : - \underline{\mu}_i \leq v_i \leq \bar{\mu}_i, i \in \pmx{1 & \dots & n_q} \right\} 
\end{equation}
\begin{equation}\label{eqn:nonlin-inclusion}
\bmx{x \\ \Phi(x)}\tr \bmx{\bar{S} & \bar{Q} \\ \star & \bar{W}} \bmx{x \\ \Phi(x)} \geq 0
\end{equation}
\end{assumption}\textcolor{red}{Please help in adding details about $\Phi$ since it's NOLCOS}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/first_scheme}
    \caption{Feedback system}
    \label{fig:first_scheme}
\end{figure}

The controller is implemented as a Multi Layer Perceptron (MLP) with $l$ layers, each with $n_{\phi_i}, i \in \left\{ 1, \dots, l\right\}$ neurons, all activation functions are saturation functions, represented by $\phi_i$. The objective is to design an Event-Triggering Mechanism (ETM) that allows to ease the computational burden associated with the computation of many non-linear activation functions providing at the same time sufficient sector conditions to handle the non-linearity of the system and the controller. The controller $\pi$ is defined by:
\begin{equation}\label{eqn:nn-equations}
  \begin{aligned}
  \hat{\omega}^{0}(k) &= x(k) \\
  \nu^{i}(k) &= W^{i} \hat{\omega}^{i - 1}(k) + b^{i}, i \in \left\{ 1, \dots, l \right\}\\
  \omega^{i}(k) &= \phi_i(\nu^i(k))\\
  \bar{u}(k) &= \phi_l(\nu^l(k))
  \end{aligned} 
\end{equation}
with $\nu^i(k) \in \RR^{n_{\phi_i}}$ the input to the $i$-th activation function $\phi^i: \RR^{n_{\phi_i}} \to \RR^{n_{\phi_i}}$, $\omega^i(k), \hat{\omega}^i(k) \in \RR^{n_{\phi_i}}$ the current and last forwarded output respectively. The weights $W^i \in \RR^{n_{\phi_i} \times n_{\phi_{i-1}}}$ and biases $b^i \in \RR^{n_{\phi_i}}$ define the affine operation of each layer. The saturation function application is to be intended element-wise and will be denoted as $\phi^i(\nu^i(k)) = \left[ \varphi(\nu^i_1(k)), \dots, \varphi(\nu^i_{n_{\phi_i}}(k)) \right]$ and $\varphi: \RR \to \RR$ is the scalar activation function assumed to be symmetric and identical for every neuron. Adopting the same notation as in \cite{css-paper} we define the controller policy in a condensed form. Denoting the augmented vectors:
\begin{equation*}
  \nu = \bmx{\nu^1 \\ \vdots \\ \nu^l}, \omega = \bmx{\omega^1 \\ \vdots \\ \omega^l}, \phi = \bmx{\phi^1(\nu^1) \\ \vdots \\ \phi^l(\nu^l)} 
\end{equation*}
with $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \RR^{n_{\phi}} \to \RR^{n_{\phi}}$ the combined non-linearity we have $\omega = \phi(\nu)$. Finally, conditions \eqref{eqn:nn-equations} can be rewritten as:
\begin{equation*}
  \bmx{u(k)\\ \nu(k)} = \overbar N \bmx{x(k) \\ \omega(k) \\ 1} 
\end{equation*}
where
\begin{equation}\label{eqn:first-N}
  \overbar N = \bmx{
    \begin{array}{c | c c c c | c}
      \0 & \0 & \dots & \0 & W^l & b^l \\ 
      \hline
      W^1 & \0 & \dots & \0 & \0 & b^1 \\
      \0 & W^2 & \dots & \0 & \0 & b^2 \\
      \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
      \0 & \0 & \dots & W^{l-1} & \0 & b^{l-1}
    \end{array}
  }
\end{equation}
Similarly as in \cite{css-extended} we implement an ETM in every layer of the controller, we add also a final one to the output that will be used to further reduce the total number of events and calls to the non-linear functions. The system and the controller are now described by figure \ref{fig:second_scheme}.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/second_scheme}
    \centering
    \caption{Feedback system subject to ETM in the controller}
    \label{fig:second_scheme}
\end{figure}
The dynamics will be rewritten as:
\begin{equation}\label{eqn:system-dynamics}
  x^{+} = A x + B u + C \Phi(E x) + D d
\end{equation}
with $u \in \RR^{n_u}$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \eqref{eqn:first-N} as:
\begin{equation}\label{eqn:last-N}
  N = \bmx{
    \begin{array}{c | c c c c | c}
      \0 & \0 & \dots & \0 & \Id & 0 \\ 
      \hline
      W^1 & \0 & \dots & \0 & \0 & b^1 \\
      \0 & W^2 & \dots & \0 & \0 & b^2 \\
      \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
      \0 & \0 & \dots & W^l & \0 & b^l
    \end{array}
  } = \bmx{
    \begin{array}{c | c | c}
    N_{u x} & N_{u \omega} & N_{u b} \\
    \hline
    N_{\nu x} & N_{\nu \omega} & N_{\nu b}
    \end{array} 
  }
\end{equation}

The system is compliant with the assumptions of \citep[Lemma 2]{css-extended} that allow us to define the following expressions regarding the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller given $x_*$, defining the matrices:
\begin{equation}
    \begin{aligned}
         &R = (\Id - N_{\nu \omega})^{-1}\\
         &R_{\omega} = N_{u x} + N_{u \omega} R N_{\nu x}\\
         &R_b = N_{u \omega} R N_{\nu b} + N_{u b}
    \end{aligned}
\end{equation}

With $R$ always invertible due to $N_{\nu \omega}$ always invertible due to its lower triangular structure. We find the equilibrium of the state by the use of a numerical solver that minimizes the following expression since it is no longer explicit due to the non-linearity $\Phi$. Given a constant disturbance $\bar{d}$:
\begin{equation}
  x_* = \min_{x} \left[(A + B R_{\omega} - \Id) x + C \Phi(E x) + D \bar{d} + B R_b \right]
\end{equation}
\begin{equation}\label{eqn:equilibrium}
  \bmx{u_* \\ \nu_*} = N \bmx{x_* \\ \omega_* \\ 1}; \ \  
  \begin{aligned}
    \omega_* &= \nu_*\\
    \nu_* &= R N_{\nu x} x_* + R N_{\nu b}\\
    u_* &= R_{\omega} x_* + R_b
  \end{aligned}
\end{equation}
All the non-linearity of the system and the controller are selected as saturation functions. With respect to \citep[Lemma 3]{css-extended}, we declare the local sector conditions that hold for every state included in the following set:

\begin{remark}\label{rem:sec-set}
\emph{For $i \in \left\{ 1, \dots, l \right\}$ given a matrix $G^i \in \RR^{n_{\phi_i} \times n_x}$, if $x$ belongs to the set}
\begin{equation}\label{eqn:inclusion-set}
S = \left\{ x \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}^i_* \leq G^i (x - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
\end{equation}

\emph{Then the following quadratic constraint holds for any diagonal positive definite matrix $T^i \in \RR^{n_{\phi_i} \times n_{\phi_i}}$:}
\begin{equation}\label{eqn:first_sec_cond}
  \left[ \nu^i - \omega^i \right]\tr T^i \left[ G^i (x - x_*) - (\omega^i - \omega^i_*) \right] \leq 0
\end{equation}\end{remark}
Where $\bar{\nu}^i$ is the saturation level of the $i$-th layer.

In order to have a lighter notation the dead-zone vector  will be defined as $\psi^i = \nu^i - \omega^i$ and $\hat{\psi}^i = \nu^i - \hat{\omega}^i$. Furthermore, all following discussions will take into account incremental variables with respect to their equilibrium, denoted with a tilde, i.e. $\tilde{x} = x - x_*$. Considering \eqref{eqn:equilibrium} we have $\psi_* = \nu_* - \omega_* = 0$ hence $\tilde{\psi} = \psi$. We define $\xi^i = \bmx{\tilde{x}\tr & \tilde{\psi}_i\tr & \tilde{\nu}_i\tr}\tr$ and $\hat{\xi}^i = \bmx{\tilde{x}\tr & \hat{\tilde{\psi}}_i\tr & \tilde{\nu}_i\tr}\tr$, then the overall sector conditions \eqref{eqn:first_sec_cond} can be rewritten with no loss of generality $\forall i \in \pmx{1 & \dots & l}$ as:
\begin{equation}\label{eqn:sec_cond}
\begin{aligned}
    &({\tilde{\psi}^i})\tr T^i \bmx{G^i \tilde{x} + \tilde{\psi}^i - \tilde{\nu}^i} = {\xi^i}\tr \bmx{0 & 0 & 0\\
    T^i G^i & T^1 & -T^1\\
    0 & 0 & 0} \xi^i = \\
    &= {\xi^i}\tr \Omega^i \xi^i \leq 0
\end{aligned}
\end{equation}

\begin{remark}\label{rem:sec-cond}
Note how ensuring ${\xi^i}\tr \Omega^i \xi^i < 0$ guarantees that the sector condition \eqref{eqn:sec_cond} holds.
\end{remark}

We can express now the new interdependency between the variables of interest:
\begin{equation}\label{eqn:incremental}
  \begin{aligned}
    \tilde{u} &= R_{\omega} \tilde{x} - N_{u \omega} R \tilde{\psi}\\
    \tilde{\nu} &= R N_{\nu x} \tilde{x} + (\Id - R) \tilde{\psi}
  \end{aligned}
\end{equation}

@@

Taking inspiration from \citep[Proposition 1, Lemma 3]{css-extended} we design the ETM for the output of each layer of the controller and the final output.

\subsection{Event-triggering mechanism}
An ETM works by triggering an event that will update the current layer and propagate its value across the network. 
In the precedent work \cite{css-extended} the triggering function is defined on the basis of the saturation sector conditions in the form \eqref{eqn:sec_cond}. The ETM was defined as:
\begin{equation}
  \hat{\omega}^{i}(k) = \begin{cases}
    \phi^i(\nu^i(k)) & \text{if } {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i} > 0\\
    \hat{\omega}^{i}(k-1) & \text{otherwise}
  \end{cases}
\end{equation}

This approach acts as a static ETM triggering an event immediately after the sector conditions are violated. Inspired by the work of \cite{data-driven} we define a dynamic quantity to be implemented in the ETM triggering condition. The idea is to lower the conservatism of the controller by introducing a time-varying threshold that will adapt to the current state of the system and controller reducing the overall number of events and computational burden. The new triggering function for the new \emph{dynamic-ETM} for the $i$-th layer is defined as:
\begin{equation}\label{eqn:dynamic_trig}
  \hat{\omega}^{i}(k) = \begin{cases}
    \phi^i(\nu^i(k)) & \text{if } {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i} > \rho^i \eta_i\\
    \hat{\omega}^{i}(k-1) & \text{otherwise}
  \end{cases};
\end{equation}

Where $\eta^i$ characterized by the following dynamics:
\begin{equation}\label{eqn:eta_dynamics}
  {\eta^i}^+ = \rho^i \eta^i - {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i}\quad  \forall i \in \left( 1, \dots, l \right) 
\end{equation}

We denote a condensed notation by taking into account all conditions for each layer
\begin{equation}
\begin{aligned}
   \bR &= \text{diag} \pmx{\rho^1 & \dots & \rho^l }\\
   \mathbf{\Psi} &= \bmx{ {\hat{\xi^1}}\tr \Omega^1 \hat{\xi^1} & \dots & {\hat{\xi^l}}\tr \Omega^l \hat{\xi^l}}\tr\\
   \boldsymbol{\eta} &= \bmx{ \eta^1 & \dots & \eta^l}\tr   
\end{aligned}
\end{equation}

@@

\subsection{Finsler Lemma application}
After the implementation of the dynamic threshold a further improvement of the triggering condition has been taken in consideration. By recalling the triggering condition:
$$
\hat{{\xi^i}}\tr \Omega^i \hat{\xi^i} > \rho^i \eta^i
$$
Intuitively the lower the left-hand term of the inequality the lower will be the number of triggered events. Unfortunately $\Omega^i$ structure is non-flexible due to his strict relation with the sector condition expression. The following lemma addresses the issue by designing some triggering matrices that guarantee a lower update rate for each layer by the use of Finsler's lemma:

\begin{lemma}\label{lem:finsler} \emph{Considering the triggering policy \eqref{eqn:dynamic_trig}, $R^i$ such that ${\xi^i}\tr = \!R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr\! = R^i \xi\tr$, with reference to the structure in \eqref{eqn:sec_cond} and assuming \eqref{eqn:incremental} holds for some $\underline{\xi}, \bar{\xi} \in \RR^{n_x + 2n_{\phi}}$ such that $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists X^i \in\! \RR^{(nx + 2n_{\phi_i}) \times (nx + 2n_{\phi_i})},$ $N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ a diagonal matrix $T^i > 0 \in \RR^{n_{\phi_i} \times n_{\phi_i}}, G^i \in \RR^{n_{\phi_i} \times n_x}$ such that:
\begin{equation}\label{eqn:finsler}
     {R^i}\tr\! He\left\{X^i - \Omega^i\right\}\! R^i + He\left\{ \smallpmat{N^i_1\\ N^i_2\\ N^i_3}\! \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0
\end{equation}
Then we have ${\xi^i}\tr X^i \xi^i \leq {\xi^i}\tr \Omega^i \xi^i$}
\end{lemma}

>>>>

\title{{\color{red}\textbf{TODO}}}

@@

\address[First]{\color{red}\textbf{TODO}}
\address[Second]{\color{red}\textbf{TODO}}

\begin{abstract}
{\color{red}\textbf{TODO}}
\end{abstract}

\begin{keyword}
{\color{red}\textbf{TODO}}
\end{keyword}

@@

\section{Introduction}
{\color{red}\textbf{TODO}}\\
\emph{Notation:} $\NN, \RR^{n}, \RR^{n \times m}$ denote the set of natural non-negative integers, the set of real vectors of dimension $n$ and the set of real matrices of dimension $n \times m$, respectively. For any matrix $A$, $A\tr$ is its transpose. For any square matrix $A$, the operator $\He{ A } = A + A\tr$ is defined. $\texttt{diag}(A_1, A_2)$ is a block-diagonal matrix with block-diagonal matrices $A_1$ and $A_2$. For a partitioned matrix, the symbol $\star$ stands for symmetric blocks. $\1_n = \bmx{1 & \dots & 1}\tr \in \NN^{n \times 1}$. We identify with subscript $i$ the $i$-th element of a vector or the $i$-th row of a matrix.

\section{Problem Formulation}
We consider the nonlinear discrete-time system stabilized by a Neural Network (NN) controller, as represented in Figure \ref{fig:first_scheme} and described by:
\begin{equation}
  x^{+} = A x + B \texttt{sat}(\bar{u}) + C \Phi(E x) + D d 
\end{equation}
where the saturation function \texttt{sat} is defined as:
$$
    \texttt{sat}(\bar{u}) = \texttt{sign}(\bar{u})\text{min}(|\bar{u}|, \widehat{u})
$$
here $\widehat{u}$ represents the limit of the input signal, $x \in \mathbb{R}^{n_x}$ is the state vector, $\bar{u} \in \mathbb{R}^{n_u}$ denotes the raw output of the controller, $d \in \mathbb{R}^{n_d}$ is a constant disturbance input, and $\Phi \in \mathbb{R}^{n_q}$ characterizes the nonlinearities of the system.
\begin{assumption}\label{ass:nonlin}
We assume that $\Phi$ satisfies some local quadratic abstraction for some $\underline{\mu}, \bar{\mu} > 0 \in \RR^{n_q}, \bar{S}, \bar{Q}, \bar{W} \in \RR^{n_q \times n_q}$
\begin{equation}\label{eqn:nonlin-set}
\forall v \in \bar{S} := \left\{ v \in  \RR^{n_q} : - \underline{\mu}_i \leq v_i \leq \bar{\mu}_i, i \in \pmx{1 & \dots & n_q} \right\} 
\end{equation}
\begin{equation}\label{eqn:nonlin-inclusion}
\bmx{x \\ \Phi(x)}\tr \bmx{\bar{S} & \bar{Q} \\ \star & \bar{W}} \bmx{x \\ \Phi(x)} \geq 0
\end{equation}
\end{assumption}\textcolor{red}{Please help in adding details about $\Phi$ since it's NOLCOS}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/first_scheme}
    \caption{Feedback system}
    \label{fig:first_scheme}
\end{figure}
The controller is implemented as a Multi-Layer Perceptron (MLP) with $l$ layers, each containing $n_{\phi_i}$ neurons for $i \in {1, \dots, l}$. All activation functions are saturation functions, represented by $\phi_i$. The goal is to design an Event-Triggering Mechanism (ETM) that alleviates the computational burden associated with evaluating numerous nonlinear activation functions while simultaneously providing sufficient sector conditions to manage the nonlinearities of both the system and the controller. The controller $\pi$ is defined as:
\begin{equation}\label{eqn:nn-equations}
  \begin{aligned}
  \widehat{\omega}^{0}(k) &= x(k) \\
  \nu^{i}(k) &= W^{i} \widehat{\omega}^{i - 1}(k) + b^{i}, i \in \left\{ 1, \dots, l \right\}\\
  \omega^{i}(k) &= \phi_i(\nu^i(k))\\
  \bar{u}(k) &= \phi_l(\nu^l(k))
  \end{aligned} 
\end{equation}
$\nu^i(k) \in \mathbb{R}^{n_{\phi_i}}$ represents the input to the $i$-th activation function $\phi^i: \mathbb{R}^{n_{\phi_i}} \to \mathbb{R}^{n_{\phi_i}}$, and $\omega^i(k), \widehat{\omega}^i(k) \in \mathbb{R}^{n_{\phi_i}}$ denote the current and last forwarded outputs, respectively. The weights $W^i \in \mathbb{R}^{n_{\phi_i} \times n_{\phi_{i-1}}}$ and biases $b^i \in \mathbb{R}^{n_{\phi_i}}$ define the affine transformation of each layer. The application of the saturation function is element-wise and is denoted as $\phi^i(\nu^i(k)) = \left[ \varphi(\nu^i_1(k)), \dots, \varphi(\nu^i_{n_{\phi_i}}(k)) \right]$, where $\varphi: \mathbb{R} \to \mathbb{R}$ is a scalar activation function that is assumed to be symmetric and identical for every neuron.

Following the notation in \cite{css-paper}, the controller policy is expressed in a condensed form. Introducing the augmented vectors:
\begin{equation*}
  \nu = \bmx{\nu^1 \\ \vdots \\ \nu^l}, \omega = \bmx{\omega^1 \\ \vdots \\ \omega^l}, \phi = \bmx{\phi^1(\nu^1) \\ \vdots \\ \phi^l(\nu^l)} 
\end{equation*}
With $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \mathbb{R}^{n_{\phi}} \to \mathbb{R}^{n_{\phi}}$ representing the combined nonlinearity, we have $\omega = \phi(\nu)$. Finally, the conditions in \eqref{eqn:nn-equations} can be reformulated as:
\begin{equation*}
  \bmx{u(k)\\ \nu(k)} = \overbar N \bmx{x(k) \\ \omega(k) \\ 1} 
\end{equation*}
where
\begin{equation}\label{eqn:first-N}
  \overbar N := \bmx{
    \begin{array}{c | c c c c | c}
      \0 & \0 & \dots & \0 & W^l & b^l \\ 
      \hline
      W^1 & \0 & \dots & \0 & \0 & b^1 \\
      \0 & W^2 & \dots & \0 & \0 & b^2 \\
      \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
      \0 & \0 & \dots & W^{l-1} & \0 & b^{l-1}
    \end{array}
  }
\end{equation}
Similarly to \cite{css-extended}, we implement an ETM in every layer of the controller, with an additional one at the output. This final ETM is designed to further reduce the total number of events and calls to the nonlinear functions. The system and controller are now represented in Figure \ref{fig:second_scheme}.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{Figures/second_scheme}
    \centering
    \caption{Feedback system subject to ETM in the controller}
    \label{fig:second_scheme}
\end{figure}

The system dynamics are rewritten as:
\begin{equation}\label{eqn:system-dynamics}
x^{+} = A x + B u + C \Phi(E x) + D d
\end{equation}
where $u \in \mathbb{R}^{n_u}$ is the output of the controller subject to the ETM. This modification requires redefining the matrix $\overbar N$, originally defined in the controller policy \eqref{eqn:first-N}, as:
\begin{equation}\label{eqn:last-N}
  N := \bmx{
    \begin{array}{c | c c c c | c}
      \0 & \0 & \dots & \0 & \Id & 0 \\ 
      \hline
      W^1 & \0 & \dots & \0 & \0 & b^1 \\
      \0 & W^2 & \dots & \0 & \0 & b^2 \\
      \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
      \0 & \0 & \dots & W^l & \0 & b^l
    \end{array}
  } = \bmx{
    \begin{array}{c | c | c}
    N_{u x} & N_{u \omega} & N_{u b} \\
    \hline
    N_{\nu x} & N_{\nu \omega} & N_{\nu b}
    \end{array} 
  }
\end{equation}
The system adheres to the assumptions outlined in \citep[Lemma 2]{css-extended}, which allow us to derive the following expressions for the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller, given $x_*$. The matrices are defined as:
\begin{equation}
    \begin{aligned}
         &R = (\Id - N_{\nu \omega})^{-1}\\
         &R_{\omega} = N_{u x} + N_{u \omega} R N_{\nu x}\\
         &R_b = N_{u \omega} R N_{\nu b} + N_{u b}
    \end{aligned}
\end{equation}
Given the lower triangular structure of $N_{\nu \omega}$, the expression for $R$ is always invertible and exists. The equilibrium of the state is found using a numerical solver that minimizes the following expression, as it is no longer explicit due to the nonlinearity $\Phi$. Given a constant disturbance $\bar{d}$:
\begin{equation}
  x_* = \min_{x} \left[(A + B R_{\omega} - \Id) x + C \Phi(E x) + D \bar{d} + B R_b \right]
\end{equation}
\begin{equation}\label{eqn:equilibrium}
  \bmx{u_* \\ \nu_*} = N \bmx{x_* \\ \omega_* \\ 1}; \ \  
  \begin{aligned}
    \omega_* &= \nu_*\\
    \nu_* &= R N_{\nu x} x_* + R N_{\nu b}\\
    u_* &= R_{\omega} x_* + R_b
  \end{aligned}
\end{equation}
All the nonlinearities of the system and the controller are chosen to be saturation functions. In accordance with \citep[Lemma 3]{css-extended}, we state the local sector conditions that hold for every state within the following set:
\begin{remark}\label{rem:sec-set}
\emph{For $i \in \left\{ 1, \dots, l \right\}$ given a matrix $G^i \in \RR^{n_{\phi_i} \times n_x}$, if $x$ belongs to the set}
\begin{equation}\label{eqn:inclusion-set}
S := \left\{ x \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}^i_* \leq G^i (x - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
\end{equation}

\emph{Then the following quadratic constraint holds for any diagonal positive definite matrix $T^i \in \RR^{n_{\phi_i} \times n_{\phi_i}}$:}
\begin{equation}\label{eqn:first_sec_cond}
  \left[ \nu^i - \omega^i \right]\tr T^i \left[ G^i (x - x_*) - (\omega^i - \omega^i_*) \right] \leq 0
\end{equation}\end{remark}
Where $\bar{\nu}^i$ represents the saturation level of the $i$-th layer.

For ease of reference, the dead-zone vector is defined as $\psi^i = \nu^i - \omega^i$ and $\widehat{\psi}^i = \nu^i - \widehat{\omega}^i$. Additionally, all subsequent discussions will use incremental variables relative to their equilibrium, denoted with a tilde, i.e., $\tilde{x} = x - x_*$. Considering \eqref{eqn:equilibrium}, we have $\psi_* = \nu_* - \omega_* = 0$, hence $\tilde{\psi} = \psi$. We define $\xi^i = \begin{bmatrix} \tilde{x}^\top & \tilde{\psi}_i^\top & \tilde{\nu}_i^\top \end{bmatrix}^\top$ and $\widehat{\xi}^i = \begin{bmatrix} \tilde{x}^\top & \widehat{\tilde{\psi}}_i^\top & \tilde{\nu}_i^\top \end{bmatrix}^\top$. Then, the overall sector conditions \eqref{eqn:first_sec_cond} can be rewritten, with no loss of generality, for all $i \in {1, \dots, l}$ as:
\begin{equation}\label{eqn:sec_cond}
\begin{aligned}
    &({\tilde{\psi}^i})\tr T^i \bmx{G^i \tilde{x} + \tilde{\psi}^i - \tilde{\nu}^i} = {\xi^i}\tr \bmx{0 & 0 & 0\\
    T^i G^i & T^1 & -T^1\\
    0 & 0 & 0} \xi^i = \\
    &= {\xi^i}\tr \Omega^i \xi^i \leq 0
\end{aligned}
\end{equation}

\begin{remark}\label{rem:sec-cond}
Note that ensuring ${\xi^i}^\top \Omega^i \xi^i < 0$ guarantees that the sector condition \eqref{eqn:sec_cond} holds.
\end{remark}

We can now express the new interdependency between the variables of interest as follows:
\begin{equation}\label{eqn:incremental}
  \begin{aligned}
    \tilde{u} &= R_{\omega} \tilde{x} - N_{u \omega} R \tilde{\psi}\\
    \tilde{\nu} &= R N_{\nu x} \tilde{x} + (\Id - R) \tilde{\psi}
  \end{aligned}
\end{equation}

@@

Inspired by \citep[Proposition 1, Lemma 3]{css-extended}, we design the ETM for the output of each layer of the controller and the final output. This is achieved by first incorporating a dynamic threshold and then refining the triggering function using Finsler's lemma.

\subsection{Event-triggering mechanism}
An ETM operates by triggering an event to update the current layer and propagate its value throughout the network. In the previous work \cite{css-extended}, the triggering function was designed based on the saturation sector conditions in the form of \eqref{eqn:sec_cond}. The ETM was defined as:
\begin{equation}
  \widehat{\omega}^{i}(k) = \begin{cases}
    \phi^i(\nu^i(k)) & \text{if } {\widehat{\xi^i}}\tr \Omega^i \widehat{\xi^i} > 0\\
    \widehat{\omega}^{i}(k-1) & \text{otherwise}
  \end{cases}
\end{equation}
This approach functions as a static ETM, triggering an event immediately when the sector conditions are violated. Building upon the concepts from \cite{data-driven}, we propose a dynamic quantity to be incorporated into the ETM triggering condition. The goal is to reduce the conservatism of the controller by introducing a time-varying threshold that adapts to the current state of the system and controller, thereby decreasing the overall number of events and computational burden. The new triggering function for the proposed \emph{dynamic-ETM} for the $i$-th layer is defined as:
\begin{equation}\label{eqn:dynamic_trig}
  \widehat{\omega}^{i}(k) = \begin{cases}
    \phi^i(\nu^i(k)) & \text{if } {\widehat{\xi^i}}\tr \Omega^i \widehat{\xi^i} > \rho^i \eta_i\\
    \widehat{\omega}^{i}(k-1) & \text{otherwise}
  \end{cases};
\end{equation}

Where $\eta^i$ is characterized by the following dynamics:
\begin{equation}\label{eqn:eta_dynamics}
  {\eta^i}^+ = \rho^i \eta^i - {\widehat{\xi^i}}\tr \Omega^i \widehat{\xi^i}\quad  \forall i \in \left( 1, \dots, l \right) 
\end{equation}

We introduce a condensed notation that incorporates all conditions across each layer.
\begin{equation}
\begin{aligned}
   \bR &= \text{diag} \pmx{\rho^1 & \dots & \rho^l }\\
   \mathbf{\Psi} &= \bmx{ {\widehat{\xi^1}}\tr \Omega^1 \widehat{\xi^1} & \dots & {\widehat{\xi^l}}\tr \Omega^l \widehat{\xi^l}}\tr\\
   \boldsymbol{\eta} &= \bmx{ \eta^1 & \dots & \eta^l}\tr   
\end{aligned}
\end{equation}

@@

\subsection{Finsler Lemma application}
Following the implementation of the dynamic threshold, a further refinement of the triggering condition has been considered. Recalling the triggering condition:
$$
\widehat{{\xi^i}}\tr \Omega^i \widehat{\xi^i} > \rho^i \eta^i
$$
Intuitively, the smaller the left-hand side of the inequality, the fewer events will be triggered. However, the structure of $\Omega^i$ is inflexible due to its strict relationship with the sector condition expression. The following lemma addresses this limitation by introducing triggering matrices designed to ensure a lower update rate for each layer, leveraging Finsler's lemma:

\begin{lemma}\label{lem:finsler} \emph{Considering the triggering policy \eqref{eqn:dynamic_trig}, $R^i$ such that ${\xi^i}\tr = \!R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr\! = R^i \xi\tr$, with reference to the structure in \eqref{eqn:sec_cond} and assuming \eqref{eqn:incremental} holds for some $\underline{\xi}, \bar{\xi} \in \RR^{n_x + 2n_{\phi}}$ such that $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists X^i \in\! \RR^{(nx + 2n_{\phi_i}) \times (nx + 2n_{\phi_i})},$ $N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ a diagonal matrix $T^i \succ 0 \in \RR^{n_{\phi_i} \times n_{\phi_i}}, G^i \in \RR^{n_{\phi_i} \times n_x}$ such that:
\begin{equation}\label{eqn:finsler}
     {R^i}\tr\! He\left\{X^i - \Omega^i\right\}\! R^i + He\left\{ \smallpmat{N^i_1\\ N^i_2\\ N^i_3}\! \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \preceq 0
\end{equation}
Then we have ${\xi^i}\tr X^i \xi^i \leq {\xi^i}\tr \Omega^i \xi^i$}
\end{lemma}

\begin{proof} By pre- and post-multiplying for $\xi$ the previous expression we end up with
$$
    {\xi^i}\tr\! He \!\left\{\!X^i \!-\! \Omega^i\right\} \xi^i + He\!\left\{ \xi\tr\! \smallpmat{N^i_1\\ N^i_2\\ N^i_3}\! \smallpmat{R N_{\nu x} & \Id - R & -\Id} \xi\! \right\}\! \leq 0
$$
By expanding the term $\pmx{R N_{\nu x} & \Id - R & -\Id} \xi$ we end up with condition \eqref{eqn:incremental}, meaning that the expression is reduced to
$$
    {\xi^i}\tr \He{X^i - \Omega^i} \xi^i \leq 0
$$
By further expanding the expression and noting that it eventually reduces to a scalar, for which the transpose is equivalent to itself, we obtain:
\begin{equation*}
\begin{aligned}
    &{\xi^i}\tr (X^i + {X^i}\tr) \xi^i \leq {\xi^i}\tr (\Omega^i + {\Omega^i}\tr) \xi^i\\
    &2 {\xi^i}\tr X^i\xi^i \leq 2 {\xi^i}\tr \Omega^i \xi^i\\
    & {\xi^i}\tr X^i \xi^i \leq {\xi^i}\tr \Omega^i \xi^i
\end{aligned}
\end{equation*}
\end{proof}
